A new tidy data structure to support exploration and modeling of temporal data

### 2019-01-19 main.Rmd
done with energy

Sensors have been put up to collect data for the project of smart city across major cities in Australia. One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter]. Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** in this paper. Half-hourly general supply in kwH have been recored for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total. Customer's demographic data provides explanatory variables other than time in a different data table. Two data tables might be joined to explore different sources that contribute to daily electricity use when needed.

@@

During a power outage, electricity usage for some households may become unavailable, thus resulting in implicit missing values in the database. Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset. It would be interesting to explore these missing patterns as part of preliminary analysis. Since the smart meter has been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for the households, instead of full lengths across the households. Figure \@ref(fig:count-gaps) reveals the gaps for the top 49 households and the rest collapsed into a single batch arranged in rows from high to low in tallies. Missing values can be found at any time point during the entire time span. A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot. On the other hand, the majority suffer occasional breakdown with more frequent happenings in January.

@@

(ref:count-gaps-cap) Time gap plots for the top 49 customers with most implicit missing values and the rest grouped into the one. Each cross represents the missing in time and a line between two dots shows continuous missingness over time. Each row corresponds to one customer describes the incidences of time gaps.

Aggregation across all individuals helps to sketch a big picture about if the representative behaviors change over time, laid out in the calendar display shown as Figure \@ref(fig:calendar-plot). Each glyph represents the daily pattern of residential electricity usage every thirty minutes with higher consumption in the time when human are more active. The daily snapshots diverge in traces, depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the hot days with daily average temperature greater than 25 degrees. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This also illustrates how the tsibble data can easily integrate with other tools and graphics.

@@

(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized in the calendar format. Distinctive patterns between summer and winter months occur to daily residential energy consumption. Air conditioning drives the peak in the late afternoon in January and December, however, peaks in the morning and evening are more pronounced in the winter months due to heating use.

>>>>

Sensors have installed in households across major cities in Australia to collect data for the smart city project. One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter]. Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study. Half-hourly general supply in kwH have been recored for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total. Customers' demographic data provides explanatory variables other than time in a different data table. Two data tables might be joined to explore different sources that contribute to daily electricity use when needed.

@@

During a power outage, electricity usage for some households may become unavailable, thus resulting in implicit missing values in the database. Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset. It would be interesting to explore these missing patterns as part of a preliminary analysis. Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household. Figure \@ref(fig:count-gaps) shows the gaps for the top 49 households arranged in rows from high to low in tallies. (The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span. A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot. On the other hand, the majority suffer occasional outages with more frequent occurrence in January.

@@

(ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at top. Each cross represents a missing in time and a line between two dots shows continuous missingness over time. Each row corresponds to one customer. Missing values occur at various times, with perhaps more in January.

Aggregation across all individuals helps to sketch a big picture about the representative behaviors change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)). Each glyph represents the daily pattern of residential electricity usage every thirty minutes.  Higher consumption is indicated by higher values, and typically occurs in daylight hours. Colour indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the hot days with daily average temperature greater than 25 degrees. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.

@@

(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized into calendar format, with colour indicating hot days. Energy use of hot days tends to be higher, suggesting air conditioner use. Days in the winter months have a double peak suggesting heater use.
